An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems

### 2014-04-29 paper/paper.tex
Paper: incorporate suggestions from mihaly into intro/bg

Most backup systems that are in use today are centralized, meaning that they rely on some central entity to store their data. This entity is typically a company that offers data backup as a service, such as Dropbox or Google. While these solutions are somewhat convenient, since the service is guaranteed to almost always be online, there are several problems with this approach. The amount of data that users can backup is severely limited, and obtaining more space costs money. In a time where owning terabytes of data is common, cloud storage meant for everyday use is not a feasible backup solution. Dedicated backup solutions can offer an adequate amount of space, but often require expensive recurring payments for their service. In addition to this, users must trust the entity to which they send their data to both remain online and respect their privacy. This trust is not always possible.

@@

\chapter{Background} \label{chap:Background}
\section{Centralized Backup Systems}
Several popular backup systems currently in use have a centralized system design. In these systems, there are many clients that all back up their data to a central location controlled by a single entity. Common elements of these systems include a fixed amount of data storage per person, with the option to obtain more for a fee. Three of the more popular systems are Google Drive, Dropbox, which has over 200 million users \cite{dropboxusers}, and OneDrive, with over 250 million users \cite{onedriveusers}.

@@

Finally, there are often restrictions on what can be stored using centralized backup systems. This includes both the size of individual files and the amount of total data. More space is gained by paying a monthly or yearly fee, but even then the amount of space is capped at a specific size. For users who need to back up a large amount of large files, these systems are not very useful.

@@

Peer to Peer (P2P) Backup Systems are applications that allow users to back up data without the use of a central entity. Instead, each user offers some of his or her storage space in order to use the space of others as backup locations. We start this section by introducing peer to peer networks, and then discuss the technical components of existing P2P backup systems.

@@

A peer to peer network is a type of distributed network model in which participants form direct connections to each other. This is in contrast to the centralized client-server network model, in which all participating clients connect to a central server to carry out their task. Client-server network models are used on the World Wide Web: the clients are users' web browsers and the server is the website being visited. In peer to peer networks, each participant, or "peer", functions as both a client and a server; peers initiating a request take on the role of the client, and peers answering the request take on the role of the server.

@@

To find other peers, a peer in a structured overlay network will consult its routing table. This table is initially populated upon joining the network, and is kept up to date as peers join and leave. When a joining peer first contacts a bootstrap peer, it is unlikely that the bootstrap peer has the joining peer's direct neighbors in its routing table. This is because the routing table is designed to include many peers that are close, and only a few peers that are far. A joining peer, by asking the bootstrap node, will learn of the peers in the bootstrap peer's routing table that the joining node is close to. The joining peer will then contact these peers, ask for the same information, and learn of peers it is even closer to. This process repeats until the joining node learns its direct neighbors \cite{p2pSurvey}.

@@

\bibitem{dropboxusers} https://www.dropbox.com/news/company-info
\bibitem{oneddriveusers} http://blog.onedrive.com/over-250m-people-using-skydrive/
\bibitem{ChurnResilient} S. Legtchenko, S. Monnet, P. Sens, G. Muller, Churn-Resilient Replication Strategy for Peer-to-Peer Distributed Hash-Tables
\bibitem{Kademlia} P. Maymounkov, D. Mazi\`eres, Kademlia: A Peer-to-Peer Information System Based on the XOR Metric
\bibitem{pStore} C. Batten, K. Barr, A. Saraf, S. Trepetin, pStore: A Secure Peer-to-Peer Backup System
\bibitem{PeerStore} M. Landers, H. Zhang, K. Tan, PeerStore: Better Performance by Relaxing in Peer-to-Peer Backup
\bibitem{StorageSearchP2PNetworks} J. Augustine, A. Molla, E. Morsy, G. Pandurangan, Storage and Search in Dynamic Peer-to-Peer Networks
\bibitem{p2pSurvey} A. Passarella, A survey on content-centric technologies for the current Internet: CDN and P2P solutions
\bibitem{BTSyncFAQ} http://www.bittorrent.com/help/faq/sync
\bibitem{btsynctech} http://www.bittorrent.com/sync/technology
\bibitem{btsyncuserguide} http://btsync.s3-website-us-east-1.amazonaws.com/BitTorrentSyncUserGuide.pdf
\bibitem{btsyncapi} http://www.bittorrent.com/sync/developers/api
\bibitem{cryptoDef} Rivest, Ronald L. (1990). "Cryptology". In J. Van Leeuwen. Handbook of Theoretical Computer Science 1. Elsevier.
\bibitem{bittorrentDHT} http://www.bittorrent.org/beps/bep\_0005.html
\bibitem{bittorrentProtocol} http://www.bittorrent.org/beps/bep\_0003.html
\bibitem{syslogRFC1} C. Lonvick, "The BSD syslog Protocol". https://tools.ietf.org/html/rfc3164, August 2001.
\bibitem{syslogRFC2} R. Gerhards, "The syslog Protocol". https://tools.ietf.org/html/rfc5424, March 2009.
\bibitem{skypeSupernode} S. Guha, N. Daswani, R. Jain, "An Experimental Study of the Skype Peer-to-Peer VoIP System". http://saikat.guha.cc/pub/iptps06-skype/ .
\bibitem{kademliaDHT} P. Maymounkov, D. Mazi\`eres, Kademlia: A Peer-to-peer Information System Based on the XOR Metric, IPTPS '01 Revised Papers from the First International Workshop on Peer-to-Peer Systems (2002), 53-65.

>>>>

Most backup systems that are in use today are centralized, meaning that they rely on some central entity to store their data. This entity is typically a company that offers data backup as a service, such as Dropbox or Google. While these solutions are somewhat convenient, since the service is guaranteed to almost always be online, there are several problems with this approach. The amount of data that users can backup is severely limited, and obtaining more space costs money. In a time where owning terabytes of data is common, general purpose cloud storage meant for everyday use is not a feasible backup solution. Dedicated backup solutions can offer an adequate amount of space, but often require expensive recurring payments for their service. In addition to this, users must trust the entity to which they send their data to both remain online and respect their privacy. This trust is not always possible.

@@

\chapter{Background} \label{chap:Background}
\section{Centralized Backup Systems}
Several popular backup systems currently in use have a centralized system design. In these systems, there are many clients that all back up their data to a central location controlled by a single entity. Examples of these include general purpose cloud services and dedicated backup services.

Common elements of general purpose cloud services include a very limited amount of data storage per person\footnote{On the order of gigabytes to tens of gigabytes.}, with the option to obtain more for a fee. Three of the more popular systems are Google Drive, Dropbox, which has over 200 million users \cite{dropboxusers}, and OneDrive, with over 250 million users \cite{onedriveusers}.

Dedicated backup services have a relaxed data storage limit\footnote{On the order of terabytes, though it is not uncommon to see unlimited data storage offered.}, but require recurring payments to use. Two popular dedicated backup services are Backblaze and Bitcasa.

@@

Finally, in general purpose cloud services, there are often restrictions on what can be stored. This includes both the size of individual files and the amount of total data. More space is gained by paying a monthly or yearly fee, but even then the amount of space is capped at a specific size. For users who need to back up a large amount of large files, these systems are not very useful. Dedicated backup services have less limitations on the amount of available storage, but generally cost more as a result.

@@

Peer to Peer (P2P) Backup Systems do not require users to back up data to a central entity. Instead, each user offers some of his or her storage space in order to use the space of others as backup locations.\footnote{A P2P backup system can use a central entity to store metadata, such as the location of backups, and still be considered P2P, albeit not completely decentralized. It is the fact that the \emph{data} is being stored on peers that makes it P2P.} We start this section by introducing peer to peer networks, and then discuss the technical components of existing P2P backup systems.

@@

A peer to peer network is a type of distributed network model in which participants form direct connections to each other. This is in contrast to the centralized client-server network model, in which all participating clients connect to a central server to carry out their task. Client-server network models are used on the World Wide Web: the clients are users' web browsers and the server is the website being visited. In peer to peer networks, each participant, or "peer", functions as both a client and a server; peers initiating a request take on the role of the client, and peers answering the request take on the role of the server. It is common for P2P networks to make use of a tracker to help manage the organization of peers. To become completely decentralized, P2P networks could use a distributed hash table\footnote{Distributed hash tables are discussed in Section \ref{sec:backgroundDHT}.} instead.

@@

To find other peers, a peer in a structured overlay network will consult its routing table. This table is initially populated upon joining the network, and is kept up to date as peers join and leave. When a joining peer first contacts a bootstrap peer, it is unlikely that the bootstrap peer has the joining peer's direct neighbors\footnote{Given the structure of the overlay network and the joining peer's ID, there is a defined set of peers in the network that are the joining peer's direct neighbors. We are discussing how the joining peer learns who is in this set.} in its routing table. This is because the routing table is designed to include many peers that are close, and only a few peers that are far. A joining peer, by asking the bootstrap node, will learn of the peers in the bootstrap peer's routing table that the joining node is close to. The joining peer will then contact these peers, ask for the same information, and learn of peers it is even closer to. This process repeats until the joining node learns its direct neighbors \cite{p2pSurvey}.

@@

\bibitem{dropboxusers} https://www.dropbox.com/news/company-info
\bibitem{onedriveusers} http://blog.onedrive.com/over-250m-people-using-skydrive/
\bibitem{ChurnResilient} S. Legtchenko, S. Monnet, P. Sens, G. Muller, Churn-Resilient Replication Strategy for Peer-to-Peer Distributed Hash-Tables
\bibitem{Kademlia} P. Maymounkov, D. Mazi\`eres, Kademlia: A Peer-to-Peer Information System Based on the XOR Metric
\bibitem{pStore} C. Batten, K. Barr, A. Saraf, S. Trepetin, pStore: A Secure Peer-to-Peer Backup System
\bibitem{PeerStore} M. Landers, H. Zhang, K. Tan, PeerStore: Better Performance by Relaxing in Peer-to-Peer Backup
\bibitem{StorageSearchP2PNetworks} J. Augustine, A. Molla, E. Morsy, G. Pandurangan, Storage and Search in Dynamic Peer-to-Peer Networks
\bibitem{p2pSurvey} A. Passarella, A survey on content-centric technologies for the current Internet: CDN and P2P solutions
\bibitem{BTSyncFAQ} http://www.bittorrent.com/help/faq/sync
\bibitem{btsynctech} http://www.bittorrent.com/sync/technology
\bibitem{btsyncuserguide} http://btsync.s3-website-us-east-1.amazonaws.com/BitTorrentSyncUserGuide.pdf
\bibitem{btsyncapi} http://www.bittorrent.com/sync/developers/api
\bibitem{cryptoDef} Rivest, Ronald L. (1990). "Cryptology". In J. Van Leeuwen. Handbook of Theoretical Computer Science 1. Elsevier.
\bibitem{bittorrentDHT} http://www.bittorrent.org/beps/bep\_0005.html
\bibitem{bittorrentProtocol} http://www.bittorrent.org/beps/bep\_0003.html
\bibitem{syslogRFC1} C. Lonvick, "The BSD syslog Protocol". https://tools.ietf.org/html/rfc3164, August 2001.
\bibitem{syslogRFC2} R. Gerhards, "The syslog Protocol". https://tools.ietf.org/html/rfc5424, March 2009.
\bibitem{skypeSupernode} S. Guha, N. Daswani, R. Jain, "An Experimental Study of the Skype Peer-to-Peer VoIP System". http://saikat.guha.cc/pub/iptps06-skype/ .
\bibitem{kademliaDHT} P. Maymounkov, D. Mazi\`eres, Kademlia: A Peer-to-peer Information System Based on the XOR Metric, IPTPS '01 Revised Papers from the First International Workshop on Peer-to-Peer Systems (2002), 53-65.
